\section*{Acknowledgments} 
\addcontentsline{toc}{chapter}{Acknowledgments}

It has been an amazing journey to have completed a Ph.D degree in Physics. 
During this journey, I have met a lot of great friends, collaborators and
mentors. I would like to take the chance to thank them one by one.  

First, I am very grateful to have met Prof. Benjamin Wandelt,  
my undergraduate advisor, who has always been a role model for me. 
He is my inspiration for learning about the use of Bayesian statistics 
to make inference from astrophysical data. I enjoyed our project so much that I   
decided to study observational cosmology.  

During my degree, I have enjoyed the company of my great friends and classmates, Dustin (and Kayla) 
Burns, Henry Stoltenberg, Austin Peng, Xiao Rui, Qin Qin, 
Brent Follin, Alison Mansheim, 
Austin Hoag, and Liam Damewood, among many others. I will not forget the
classes, study sessions and thesis-writing sessions 
that we have had together.
I am also very lucky to have two very kind and considerate housemates, Nancy Xuan He
and Hanmei Sun, who have taken great care of me.  
 
I am also glad to have crossed paths with many 
extraordinary physicists, statisticians and computer scientists. 
I thank Will Dawson, Michael Schneider, Debbie Bard and Annalisa Pilleprich for the intellectual
discussions as collaborators and the learning opportunities that they have
given me. 
I really appreciate the support that I have received from the
faculty members in Cosmology, including Prof. Tony Tyson, 
Prof. Chris Fassnacht and Prof. Lori Lubin. 
I am also extremely grateful to Prof. Thomas Lee and Prof. Duncan Temple Lang from the
Statistics Department. Prof. Lee has shared great statistical insights and 
advice for my research projects. I have enjoyed the lectures and workshops 
from Prof. Temple Lang on statistical computing. The skills that I have learned
from his lessons are indispensable for my research. 
 
I am very thankful to have a responsive dissertation committee, including
Prof. Maru\v{s}a Brada\v{c} and Prof. Annika Peter, who have given me 
countless valuable feedback. They have helped improve the
readability of this dissertation drastically.  
In particular, I would like to express my sincere gratitude to my Ph.D advisor,
Prof. David Wittman, who has given me a lot of freedom and support for 
me to explore my research interests over the years. 

% % Family
Finally, I would like to dedicate this thesis to my family, my mother Mei-Ying
Tsang, my sister, Chloe Yin-Ching Ng, and my father, Wing-Po Ng.
I would never have the courage to pursue my degrees in Physics if not for the
selfless, unlimited support of my family.  
They have provided me with resources and opportunities that they never had
themselves. 
I will make good use of the knowledge and skills that I have obtained as a way
of paying the favor forward. I am proud to have achieved the same degree as my 
better half and my best friend, Dr. Billy Lo. I hope to share a happy future
with Dr. Lo to thank him for encouraging me to pursue my dreams and waiting for
my graduation. 
